Introduction to Lattices Attacks\\``Cryptography''

### 2021-06-03 paper.tex
Add references RSA

\begin{definition}
    \textbf{Lineaer Independece}
\end{definition}

@@

The \textbf{Lenstra-Lenstra-Lovász} \textit{LLL} or \textit{$L^3$} is a polynomial time algorithm to find a "shorter" basis.

@@

\begin{enumerate}
    \item Factoring polynomials over the integers. For example, given $x^2 - 1$ factor it into $x + 1$ and $x - 1$.
    \item Integer Programming. This is a well-known \textbf{NP}-complete problem. Using LLL, one can obtain a polynomial time solution
          to integer programming with a fixed number of variables.
    \item Approximation to the \textbf{CVP} or \textbf{SVP}, as well as other lattice problems.
    \item Application in cryptanalysis.
\end{enumerate}

@@

% \section{Attack Knapsack}

@@

\begin{itemize}
    \item If $m < N^{\frac{1}{e}}$, then $m^e < N$ and the modulo operation is not applied and we only need to find the $e$th root of $c$ over the
        integers to find $m$.
    \item If $q = p + x$ where $x$ is small enough than it's easy to recover the factors of $N$ computing $A = \sqrt{N}$ and compute
            $p = A - x$ for different value of $x$ until $N \mod p = 0$, then we have found $q = N / p$
    \item Many more attacks can be found TODO.
\end{itemize}

@@

We want to attack a relaxed model of RSA where we know a part of the message $m$.
We start introducing the \ref{TODO} coppersmith attack and the math behind it.

@@

\subsection{Example}

We have a 100-bit \textbf{RSA} modulus

\begin{center}
    $N=0xf046522fb555a90bdc558fc93$ and $e = 3$.
\end{center}

@@

\[
    z = pad || m = 0x74686973206b65793a || m
\]

where $||$ is the concatenation. The padding is the ascii encoding of ``this key:''\\

@@

\[
    c = z^e \mod N = 0x5b603cda4b72100c6f25954fc
\]

Suppose that we don't know the factorization of $N$ and we would like to know the message $m$.
However we know the padding and that the length of $m < 2^{16}$.

@@

\[
a = 0x74686973206b65793a0000
.\]

@@

\begin{align*}
    f(x) = x^3 + 0x15d393c596142306bae0000x^2 + 0x1b53c5e184a49b39f9ad9eedbx\\
    + 0x486a5d936fb568185c8ff0506
\end{align*}

@@

\textbf{Lattice contruction}. Let the coefficients of $f$ be $f(x) = x^3 + f_2x^2 + f_1x + f_0$
and $X = 2^{16}$ be the upper bound of the size of the root $m$. We can construct the matrix

@@

\[
    \begin{split}
    v = (0x90843131bc53X^3 + 0x2736f60b1c7ba3294X^2, \\
    -0x1bec331b20625341b6d73X, 0x47336b98335c143ac912ec9e)
    \end{split}
\]

@@

\[
    \begin{split}
        g(x) = 0x90843131bc53x^3 + 0x2736f60b1c7ba3294x^2 \\
        -0x1bec331b20625341b6d73x + 0x47336b98335c143ac912ec9e
    \end{split}
\]

@@

We have $(\det B)^{\frac{1}{4}} = (X^6N^3)^{\frac{1}{4}} < \frac{N}{\sqrt{4}}$, if we solves for $X$ this will be satisfied when $X < (\frac{N}{16})^{\frac{1}{6}}$.
With the numbers we have this inequality is verified, however even if the bound of the shortest vector is larger we still have some possibilities to
find the correct root.

If we compute the root of $g(x)$ over the integers we obtain $m = 0x6162$ which is the correct result.\\

@@

To compute bigger root a bigger lattice with more polynomials generated with \ref{eq:create_pol} is needed,
this method is better described in TODO, but the principles are the same.

>>>>

\begin{definition}
    \textbf{Linear Independece}
\end{definition}

@@

The \textbf{Lenstra-Lenstra-Lovász} \textit{LLL}\cite{LLL} or \textit{$L^3$} is a polynomial time algorithm to find a ``short'' basis of a lattice.

@@

\begin{enumerate}
    \item Factoring polynomials over the integers. For example, given $x^2 - 1$ factor it into $x + 1$ and $x - 1$.
    \item Integer Programming. This is a well-known $\mathcal{NP}$-complete problem. Using \textbf{LLL}, one can obtain a polynomial time solution
          to integer programming with a fixed number of variables.
    \item Approximation to the \textbf{CVP} or \textbf{SVP}, as well as other lattice problems.
    \item Cryptanalysis (break cryptographic protocols).
\end{enumerate}

@@

@@

\begin{itemize}
    \item If $m < N^{\frac{1}{e}}$, then $m^e < N$ and the modulo operation is not applied and we only need to find the $e$th root of $c$ over the
        integers to find $m$.
    \item If $q = p + x$ where $x$ is small enough than it's easy to recover the factors of $N$ computing $A = \sqrt{N}$ and compute
            $p = A - x$ for different value of $x$ until $N \mod p = 0$, then we have found $q = N / p$.
    \item If $d$ is too small (for example to accelerate decryption) then Wiener found a way to recover the private key $d$ \cite{wiener}.
    \item Other attacks can be found on the excellent survey by Dan Boneh \cite{boneh99twentyyears}.
\end{itemize}

@@

We start introducing the main ideas of Coppersmith's attack \cite{coppersmith96} and the math behind it.
The attack can be used against \textbf{RSA} when a small $e$ is used. In 2017 the attack was used to exploit
\href{https://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2017-15361}{CVE-2017-15361} also named \textbf{ROCA}
(Return Of Coppersmith's Attack) \cite{roca}.

@@

\subsection{Recover plaintext with known MSB bits of $m$}

We will use a simplified version of the Coppersmith's attack that works when the root is approximately less than $N^{\frac{1}{6}}$.

\textbf{Problem setup}. Let $N$ be an \textbf{RSA} modulus of $97$-bit:

\begin{center}
    $N=\texttt{0x1a34992f64135aedaa0fe2bfd}$ and $e = 3$.
\end{center}

@@

\[
    z = pad || m = \texttt{0xffffffffffffffffffffff} || m
\]

where $||$ is the concatenation.

@@

\[
    c = z^e \mod N = \texttt{0x11701d5479bcdef207cba6937}
\]

Suppose that we don't know the factorization of $N$ and we would like to know the message $m$.
We know the padding and that the message $m$ is 2 bytes long, composed by only ascii letters, so $m < \texttt{0x7b7b}$.

@@

\[
    a = \texttt{0xffffffffffffffffffffff0000}
.\]

@@

\begin{align*}
    f(x) = x^3 + \texttt{0x17d7f55d0c9dc5851af54957c}x^2 + \texttt{0x3cf96fed2ea927527a2ed329}x\\
    + \texttt{0x32882e547964b3bcf75d315e}
\end{align*}

@@

\textbf{Lattice contruction}. Let the coefficients of $f$ be $f(x) = x^3 + f_2x^2 + f_1x + f_0$
and $X = \texttt{0x7b7b}$ be the upper bound of the size of the root $m$. We can construct the matrix

@@

\[
    \begin{split}
    v = (-\texttt{0xd55d67baf71d}X^3, \texttt{0x3cf3cca3200a58bf}X^2, \\
    \texttt{0x3854d44211e80f248449}X, -\texttt{0xec93bf51cf766f7b9f1e5e6})
    \end{split}
\]

@@

\[
    \begin{split}
        g(x) = -\texttt{0xd55d67baf71d}x^3 +  \texttt{0x3cf3cca3200a58bf}x^2 \\
        \texttt{0x3854d44211e80f248449}x -\texttt{0xec93bf51cf766f7b9f1e5e6}
    \end{split}
\]

@@

We have $(\det B)^{\frac{1}{4}} = (X^6N^3)^{\frac{1}{4}} < \frac{N}{\sqrt{4}}$, if we solves for $X$ this will be satisfied when
$X < (\frac{N}{16})^{\frac{1}{6}}$. We can consider the $\sqrt{4}$ an error term to have $X < N^{\frac{1}{6}}$.
With the numbers we have this inequality is verified, however even if the bound of the shortest vector is larger we still have some possibilities to
find the correct root.

If we compute the root of $g(x)$ over the integers we obtain $m = \texttt{0x4142} = $ ``AB'' which is the correct result.\\

@@

To compute bigger root a bigger lattice with more polynomials generated with \ref{eq:create_pol} is needed,
this method is better described in \cite{may2011} and a detailed implementation in sagemath by David Wong is available \cite{wong2015}.
